---
title: "Lab 06 - Ugly charts and Simpson's paradox"
author: "Olivia Zhang"
date: "02/27/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
library(mosaicData) 

staff <- read_csv("data/instructional-staff.csv")
fisheries <- read_csv("data/fisheries.csv")

#wide to long
staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))
```

### Exercise 1

```{r line-plot, fig.height=5, fig.width=10}
staff_long %>%
  ggplot(aes(
    x = year,
    y = value,
    group = faculty_type,
    color = faculty_type
  )) +
  geom_line() +
  labs(title = "Trends in Instractional Staff Employmees", 
       subtitle = "between 1975 and 2011", 
       x = "Year", 
       y = "Percentage of Hires", 
       color = "Faculty Type")
```


### Exercise 2

To make the porportion of part-time faculty pop up, I would change the color of other faculty types to black.
```{r line-plot-part-time, fig.height=5, fig.width=10}
staff_long %>%
  ggplot(aes(
    x = year,
    y = value,
    group = faculty_type,
    color = ifelse(faculty_type == "Part-Time Faculty", "red", "black")
  )) +
  geom_line() +
  labs(title = "Trends in Instractional Staff Employmees", 
       subtitle = "between 1975 and 2011", 
       x = "Year", 
       y = "Percentage of Hires", 
       color = "Faculty Type") +
  scale_color_identity() # Ensures colors are used directly
#not sure how to get the legends show up.
```


### Exercise 3

The original graphs have a few problems:
1. The y axis of the density plot is too stretch out because of the large capture and aquaculture industry in China. One way to improve this is to use take the log of the tonnage of fish farmed and captured so that the data from other countries can be seen.
2. The pie chart is kind of uninformative and conveys nothing new from the density plot. One way to improve this is to create a histogram/barplot that shows the percentage of fish capture and aquaculture for each country.

```{r density-plot, fig.height=5, fig.width=11}
#wide to long, filter out total >= 100000
fisheries_long <- fisheries %>%
  arrange(desc(total)) %>%
  filter(total >= 100000) %>%
#  select("country", "capture", "aquaculture") %>%
  pivot_longer(cols = -country, names_to = "type") %>%
  mutate(value = as.numeric(value)) 

fisheries_long %>%
  ggplot(aes(x = fct_inorder(country), y = value, fill = type)) +
#  geom_bar(stat='identity', position = "stack", width = .7) +
  geom_bar(stat='identity', width = .7) +
  facet_wrap(~ type, nrow = 3, scales = "free_y") +
  labs(
    title = "Tonnage of Total Fish Captured and Farmed in 2016",
    x = "Country",
    y = "Total Tonnage of Fish Captured and Farmed", 
    fill = "Type") +
#  scale_y_continuous(breaks = scales::breaks_extended(n = 4), labels = fisheries_long$value) +
  scale_y_continuous(labels = scales::label_number(accuracy = 1)) +
#  scale_y_continuous(
#    trans = "log",  # Apply log transformation
#    labels = scales::trans_format("log10", scales::math_format(10^.x))  # Show original y-axis values
#  ) +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1) # Tilt x-axis ticks
  ) 
```

I tried numerous ways but haven't figured out how to make the y axis clear enough to show the differences but also display the original values. 